---
title: "class05"
author: "RUNQI ZHANG"
format: pdf
---

# Our first ggplot

To use the ggplot2 package I first need to have it installed on my computer. (.)(.)

1. use the 'install.packages()' command.

ggplot is not loaded by defaul to save disk space. To manually load the package,

2. use 'library(ggplot2)' to load the ggplot2 package
```{r}
library(ggplot2)

ggplot(mpg) +
  aes(displ, hwy) +
  geom_point()
```

I can add more layers:
```{r}
ggplot(mpg) +
  aes(displ, hwy) +
  geom_point() +
  geom_smooth(se=0)
```
# Plot of gene expression data

First read the data from online.
```{r}
url <- "https://bioboot.github.io/bimm143_S20/class-material/up_down_expression.txt"
genes <- read.delim(url)
View(genes)
```


```{r}
p <- ggplot(genes) +
      aes(Condition1, Condition2, color=State) +
      geom_point() +
      geom_smooth()

p
```
Q: How mnay genes are in the dataset
```{r}
nrow(genes)
```

Q: How many genes are up/downregulated or unchanged?

```{r}
table(genes$State)
```
Q: Using your values above and 2 significant figures. What fraction of total genes is up-regulated in this dataset?
```{r}
round(table(genes$State)/nrow(genes)*100, 2)
```
# Further implementation to ggplot

```{r}
p <- ggplot(genes) +
      aes(Condition1, Condition2, color=State) +
      geom_point(alpha=.3) +
      scale_color_manual(values=c("blue","gray","red")) +
      labs(title="Gene Expression Changes Upon Drug Treatment",
           subtitle="Condition 1 VS. Condition 2",
           caption="By Runqi Zhang")

p
```

