---
title: "Lab_15"
author: "RUNQI ZHANG"
format: pdf
toc: TRUE
---

# Getting Started

```{r}
# Import vaccination data
vax <- read.csv('covid19vaccinesbyzipcode_test.csv')
head(vax)
```

```{r}
head(vax$as_of_date)
tail(vax$as_of_date)
```
## Q1. What column details the total number of people fully vaccinated?
A: persons_fully_vaccinated

## Q2. What column details the Zip code tabulation area?
A: zip_code_tabulation_area

## Q3. What is the earliest date in this dataset?
A: 2021-01-05

## Q4. What is the latest date in this dataset?
A: 2022-11-22

```{r}
skimr::skim(vax)
```
```{r}
sum (is.na(vax$persons_fully_vaccinated))

sum (is.na(vax$persons_fully_vaccinated))/nrow(vax)

```
## Q5. How many numeric columns are in this dataset? 
A: 13

## Q6. Note that there are “missing values” in the dataset. How many NA values there in the persons_fully_vaccinated column? 
A: 14921

## Q7. What percent of persons_fully_vaccinated values are missing (to 2 significant figures)? 
A: 0.085

## Q8. [Optional]: Why might this data be missing?


# Working with dates
```{r}
library(lubridate)
```
```{r}
today()
```
```{r}
# Specify that we are using the year-month-day format
vax$as_of_date <- ymd(vax$as_of_date)
```

```{r}
today() - vax$as_of_date[1]
```

```{r}
# Using the last and the first date value we can now determine how many days the dataset span?
vax$as_of_date[nrow(vax)] - vax$as_of_date[1]

# Finding how many unique dates are in the dataset
sum( !is.na( unique(vax$as_of_date) ))
```
## Q9. How many days have passed since the last update of the dataset? 
A: 691

## Q10. How many unique dates are in the dataset (i.e. how many different dates are detailed)? 
A: 99

# Working with ZIP codes
```{r}
library(terra)
library(zipcodeR)
```

```{r}
geocode_zip('92037')
```

```{r}
# Calculate the distance between the centroids of any two ZIP codes in miles, e.g.
zip_distance('92037','92109')
```


```{r}
# More usefully, we can pull census data about ZIP code areas (including median household income etc.)
reverse_zipcode(c('92037', "92109") )
```

```{r}
# Pull data for all ZIP codes in the dataset
#zipdata <- reverse_zipcode( vax$zip_code_tabulation_area )
```

# Focus on the San Diego area

```{r}
library(dplyr)

sd <- filter(vax, county == "San Diego")

nrow(sd)

sum( !is.na( unique(sd$zip_code_tabulation_area) ))
```

```{r}
sd.10 <- filter(vax, county == "San Diego" &
                age5_plus_population > 10000)

sd$zip_code_tabulation_area[which.max(sd$age12_plus_population)]

```
## Q11. How many distinct zip codes are listed for San Diego County? 
A: 107

## Q12. What San Diego County Zip code area has the largest 12 + Population in this dataset? 
A: 92154


Using dplyr select all San Diego “county” entries on “as_of_date” “2022-11-15” and use this for the following questions.
```{r}
sd.221115 <- filter(vax, county == "San Diego" &
                as_of_date == "2022-11-15")

sum(sd.221115$persons_fully_vaccinated[!is.na(sd.221115$persons_fully_vaccinated)])/sum(sd.221115$tot_population[!is.na(sd.221115$tot_population)])
```

## Q13. What is the overall average “Percent of Population Fully Vaccinated” value for all San Diego “County” as of “2022-11-15”? 
A: persons_fully_vaccinated / tot_population = 0.752562

## Q14. Using either ggplot or base R graphics make a summary figure that shows the distribution of Percent of Population Fully Vaccinated values as of “2022-11-15”?
A: see plot below
```{r}
library(ggplot2)

p1 <- ggplot(sd.221115, aes(percent_of_population_fully_vaccinated))+
        geom_histogram(bins=9)+
        xlab("Percent of Population Fully Vaccinated in a Zip Code Area")+
        ylab("count")

p1 + labs(title="Histogram of Vaccination Rates Accross San Diego County",
          subtitle="As of 2022-11-15")
  
```
# Focuse on UCSD/La Jolla
```{r}
ucsd <- filter(sd, zip_code_tabulation_area=="92037")
ucsd[1,]$age5_plus_population
```

## Q15. Using ggplot make a graph of the vaccination rate time course for the 92037 ZIP code area:
```{r}
p2 <- ggplot(ucsd) +
        aes(as_of_date,
            percent_of_population_fully_vaccinated) +
        geom_point() +
        geom_line(group=1) +
        ylim(c(0,1)) +
        labs(x="Date", y="Percent Vaccinated")
p2
```
# Comparing to similar sized areas
```{r}
# Subset to all CA areas with a population as large as 92037
vax.36 <- filter(vax, age5_plus_population > 36144 &
                as_of_date == "2022-11-15")
#head(vax.36)

sum(vax.36$persons_fully_vaccinated[!is.na(vax.36$persons_fully_vaccinated)])/sum(vax.36$tot_population[!is.na(vax.36$tot_population)])
```

## Q16. Calculate the mean “Percent of Population Fully Vaccinated” for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2022-11-15”. Add this as a straight horizontal line to your plot from above with the geom_hline() function?
```{r}
p2
p2 + geom_hline(aes(yintercept=0.7159711),
                col="RED",
                linetype="dotted")
```
```{r}
skimr::skim(vax.36)
```

## Q17. What is the 6 number summary (Min, 1st Qu., Median, Mean, 3rd Qu., and Max) of the “Percent of Population Fully Vaccinated” values for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2022-11-15”?
A: Mean: 7.172851e-01	
   Min: 3.78501e-01	
   1st Q: 6.396185e-01	
   Median: 7.15524e-01	
   3rd Q: 7.879820e-01	
   Max: 1.00000e+00	

## Q18. Using ggplot generate a histogram of this data.
A: see plot below
```{r}
p3 <- ggplot(vax.36, aes(percent_of_population_fully_vaccinated))+
        geom_histogram()+
        xlab("Percent Vaccinated")+
        ylab("Count")+
        xlim(0,1)
p3
```


```{r}
vax %>% filter(as_of_date == "2022-11-15") %>%  
  filter(zip_code_tabulation_area=="92040") %>%
  select(percent_of_population_fully_vaccinated)
```
```{r}
vax %>% filter(as_of_date == "2022-11-15") %>%  
  filter(zip_code_tabulation_area=="92109") %>%
  select(percent_of_population_fully_vaccinated)
```
## Q19. Is the 92109 and 92040 ZIP code areas above or below the average value you calculated for all these above?
A: both are below the average value previously calculated

## Q20. Finally make a time course plot of vaccination progress for all areas in the full dataset with a age5_plus_population > 36144.
```{r}
vax.36.all <- filter(vax, age5_plus_population > 36144)


ggplot(vax.36.all) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated, 
      group=zip_code_tabulation_area) +
  geom_line(alpha=0.2, color="BLUE") +
  ylim(0,1) +
  labs(x='Date', y='Percent Vaccinated',
       title='Vaccination rate across California',
       subtitle='Only areas with a populaiton above 36k are shown') +
  geom_hline(yintercept=0.7159711 , linetype='dashed')
```

Q21. How do you feel about traveling for Thanksgiving Break and meeting for in-person class afterwards?
A: I prefer Online class for next week so I can be more flexible with my schedule